Commutative Character Sheaves on smooth group schemes over finite fields

### 2016-01-27 followup.tex
Final commit for conversation with David. Finished the proof of geotypes. It needs editing and proofing, however.

\subsection{Geometrization of quasicharacters of smooth group schemes over $p$-adic fields}

@@

\subsection{Smooth models}\label{ssec:models}

@@

We now define Weil sheaves $\cs{A}^i_m$ on $G_m^i \ceq \Gr^{\OK}_m(\underline{G}^{i})$, for $i=1,\ldots ,d$, recursively, following the construction of the representations $\,^0\rho_i$, reviewed in Section~\ref{ssec:review}.
We have already defined $\cs{A}^0_m$ so that 
\[
\trFrob{\cs{A}^0_m} = \trace(\,^\circ \rho_0)_m
\]
where $\trace(\,^\circ \rho_0)_m : G_m^0(\Fq) \to \EE$ is defined by 
\[
\underline{G}^0
\]
and the canonical identification $G_m^0(\Fq) = \underline{G}^0(\OK/\mathfrak{p}_K^{m+1})$.

Suppose $\cs{A}^i_m$ on $G_m^i$ is defined; we shall construct $\cs{A}_m^{i+1}$ on $G_m^{i+1}$.

Applying the Greenberg functor $\Gr^{\OK}_m$, to \eqref{eq:pimodel} gives
\begin{equation}\label{eq:pi}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && G_m^{i}\times_{G_m} J_m^{i+1} \arrow{d} && \\
1 \arrow{r} &J_m^{i+1} \arrow{r} & G_m^i \ltimes J_m^{i+1} \arrow{d}{\pi_m^{i+1}} \arrow{r}{p_m^{i}} & G_m^i \arrow{r} & 1\\
&& G_m^{i+1} \arrow{d} && \\
&& 1 && 
\end{tikzcd}
\end{equation}
where $J_m^{i+1} \ceq \Gr^{\OK}_m(\underline{J}^{i+1})$ and $G_m^{i} \ceq \Gr^{\OK}_m(\underline{G}^{i})$
By \cite{BG-A}*{Prop 7.1}, the sequences are exact.
Consider $\cs{A}_m^{i}\otimes \cs{L}_m^{i}$ on $G_m^{i}$ and let 
\[
(p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})
\]
be the Weil sheaf on $G_m^{i}\ltimes J_m^{i+1}$ obtained by pullback along $G_m^{i}\ltimes J_m^{i+1} \to G_m^{i}$.
Consider the sheaf 
\[
\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})
\]
on $G_m^{i}\ltimes J_m^{i+1}$ and compare with \eqref{eq:tensor}.
With reference to the morphism $\pi_m^{i+1} : G_m^i \ltimes J_m^{i+1} \to G_m^{i+1}$ from \eqref{eq:pi}, define
\[
\cs{\tilde A}_m^{i+1} \ceq R(\pi_m^{i+1})_! (\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})).
\]
%Let $n = \# (G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$ and set $\cs{A}_m^{i+1} \ceq \cs{\tilde A}_m^{i+1}(-n)$ (Tate twist).
Before continuing, let us calculate: 
\begin{eqnarray*}
\trFrob{\cs{\tilde A}^{i+1}_m}(x)
&=& \sum_{y\in (\pi_m^{i+1})^{-1}(x)}  \trFrob{\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^i)}(y)\\
&=& \sum_{y\in (\pi_m^{i+1})^{-1}(x)}  \trFrob{\cs{W}_m^{i+1}}(y)\  \trFrob{\cs{A}_m^{i}}(p_m^{i+1}(y) )\ \trFrob{\cs{L}_m^i}(p_m^{i+1}(y)).
\end{eqnarray*}
Comparing with \eqref{eq:tilderho}, we see that 
\[
y \mapsto \trFrob{\cs{W}_m^{i+1}}(y)\ \trFrob{\cs{A}_m^{i}}(p_m^{i+1}(y))\ \trFrob{\cs{L}_m^i}(p_m^{i+1}(y))
\]
is precisely the function obtained by factoring the character of $\,^\circ {\tilde \rho}_{i+1}$ through $(\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK) \to (\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK/\mathfrak{p}_K^{m+1})$, using the canonical identification $(G_m^{i}\times_{G_m} J_m^{i+1})(\Fq) =  (\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK/\mathfrak{p}_K^{m+1})$. 
In particular, this is trivial on $(G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$.
It follows that 
\[
\trFrob{\cs{\tilde A}^{i+1}_m} = n \trace(\,^\circ \rho_{i+1})
\]
on $G_m^{i+1}(\Fq)$ with $n = \# (G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$.\todo{Hmm... Homomorphism?}
\todo{If $n$ is a power of $q$, then we can use Tate twists to avoid working with virtual sheaves! Is $G^{i}\cap J_m^{i+1}$ unipotent? Is so, then there's a chance.}
Let $\cs{A}_m^{i+1}$ be the \emph{virtual} Weil sheaf on $G_m^i$ given by $\cs{A}_m^{i+1} = \frac{1}{n} \cs{\tilde A}_m^{i+1}$.

This completes the inductive definition of

>>>>

\subsection{Quasicharacters of smooth group schemes over $p$-adic fields}

@@

\subsection{Review of Yu's smooth models}\label{ssec:models}

@@

We now define Weil sheaves $\cs{A}^i_m$ on $G_m^i \ceq \Gr^{\OK}_m(\underline{G}^{i})$, for $i=1,\ldots ,d$, recursively, following the construction of the representations $\,^0\rho_i$, reviewed in Section~\ref{ssec:review}.
We have already defined $\cs{A}^0_m$ so that 
\[
\trFrob{\cs{A}^0_m} = \trace(\,^\circ \rho_0)_m
\]
where $\trace(\,^\circ \rho_0)_m : G_m^0(\Fq) \to \EE$ is defined by 
\[
\begin{tikzcd}
G^0(\Fq) \arrow{rr}{\trace(\,^\circ \rho_0)} \arrow{rd} && \EE\\
& G^0_m(\Fq). \arrow{ru}[swap]{\trace(\,^\circ \rho_0)_m} & 
\end{tikzcd}
\]
and the canonical identification $G_m^0(\Fq) = \underline{G}^0(\OK/\mathfrak{p}_K^{m+1})$.
Suppose $\cs{A}^i_m$ on $G_m^i$ is defined and \[
\begin{tikzcd}
G^i(\Fq) \arrow{rr}{\trace(\,^\circ \rho_i)} \arrow{rd} && \EE\\
& G_m^i(\Fq) \arrow{ru}[swap]{\trFrob{\cs{A}_m^i}} & 
\end{tikzcd}
\]
commutes.
We shall construct $\cs{A}_m^{i+1}$ on $G_m^{i+1}$ so that 
\[
\begin{tikzcd}
G^{i+1}(\Fq) \arrow{rr}{\trace(\,^\circ \rho_{i+1})} \arrow{rd} && \EE\\
& G_m^{i+1}(\Fq)\arrow{ru}[swap]{{\trFrob{\cs{A}_m^i}} } & 
\end{tikzcd}
\]

Applying the Greenberg functor $\Gr^{\OK}_m$, to \eqref{eq:pimodel} gives
\begin{equation}\label{eq:pi}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && G_m^{i}\times_{G_m} J_m^{i+1} \arrow{d} && \\
1 \arrow{r} &J_m^{i+1} \arrow{r} & G_m^i \ltimes J_m^{i+1} \arrow{d}{\pi_m^{i+1}} \arrow{r}{p_m^{i}} & G_m^i \arrow{r} & 1\\
&& G_m^{i+1} \arrow{d} && \\
&& 1 && 
\end{tikzcd}
\end{equation}
where $J_m^{i+1} \ceq \Gr^{\OK}_m(\underline{J}^{i+1})$ and $G_m^{i} \ceq \Gr^{\OK}_m(\underline{G}^{i})$
By \cite{BG-A}*{Prop 7.1}, the sequences are exact.
Consider $\cs{A}_m^{i}\otimes \cs{L}_m^{i}$ on $G_m^{i}$ and let 
\[
(p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})
\]
be the Weil sheaf on $G_m^{i}\ltimes J_m^{i+1}$ obtained by pullback along $G_m^{i}\ltimes J_m^{i+1} \to G_m^{i}$.
Consider the sheaf 
\[
\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})
\]
on $G_m^{i}\ltimes J_m^{i+1}$ and compare with \eqref{eq:tensor}.
With reference to the morphism $\pi_m^{i+1} : G_m^i \ltimes J_m^{i+1} \to G_m^{i+1}$ from \eqref{eq:pi}, define
\[
\cs{\tilde A}_m^{i+1} \ceq R(\pi_m^{i+1})_! (\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^{i})).
\]
%Let $n = \# (G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$ and set $\cs{A}_m^{i+1} \ceq \cs{\tilde A}_m^{i+1}(-n)$ (Tate twist).
Then
\begin{eqnarray*}
\trFrob{\cs{\tilde A}^{i+1}_m}(x)
&=& \sum_{y\in (\pi_m^{i+1})^{-1}(x)}  \trFrob{\cs{W}_m^{i+1} \otimes (p_m^{i+1})^*(\cs{A}_m^{i}\otimes \cs{L}_m^i)}(y)\\
&=& \sum_{y\in (\pi_m^{i+1})^{-1}(x)}  \trFrob{\cs{W}_m^{i+1}}(y)\  \trFrob{\cs{A}_m^{i}}(p_m^{i+1}(y) )\ \trFrob{\cs{L}_m^i}(p_m^{i+1}(y)).
\end{eqnarray*}
Comparing with \eqref{eq:tilderho}, we see that 
\[
y \mapsto \trFrob{\cs{W}_m^{i+1}}(y)\ \trFrob{\cs{A}_m^{i}}(p_m^{i+1}(y))\ \trFrob{\cs{L}_m^i}(p_m^{i+1}(y))
\]
is precisely the function obtained by factoring the character of $\,^\circ {\tilde \rho}_{i+1}$ through $(\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK) \to (\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK/\mathfrak{p}_K^{m+1})$, using the canonical identification $(G_m^{i}\times_{G_m} J_m^{i+1})(\Fq) =  (\underline{G}^{i}\ltimes \underline{J}^{i+1})(\OK/\mathfrak{p}_K^{m+1})$. 
In particular, this is trivial on $(G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$.
It follows that 
\[
\trFrob{\cs{\tilde A}^{i+1}_m} = n \trace(\,^\circ \rho_{i+1})
\]
on $G_m^{i+1}(\Fq)$ with $n = \# (G_m^{i}\times_{G_m} J_m^{i+1})(\Fq)$.\todo{Hmm... Homomorphism?}
\todo{If $n$ is a power of $q$, then we can use Tate twists to avoid working with virtual sheaves! Is $G^{i}\cap J_m^{i+1}$ unipotent? Is so, then there's a chance.}
Let $\cs{A}_m^{i+1}$ be the \emph{virtual} Weil sheaf on $G_m^i$ given by $\cs{A}_m^{i+1} = \frac{1}{n} \cs{\tilde A}_m^{i+1}$.
%
This completes the inductive definition of $\cs{A}_m^i$ with the desired property.

Now set $\cs{F}_m = \cs{A}_m^{d}$.
Then $\cs{F}_m$ is a virtual Weil sheaf on $G_m = \Gr^{\OK}_m(\underline{G})$ such that
\[
\begin{tikzcd}
G(\Fq) \arrow{rr}{\trace(\,^\circ \rho)} \arrow{rd} && \EE\\
& G_m(\Fq) \arrow{ru}[swap]{\trFrob{\cs{A}_m}} & 
\end{tikzcd}
\]
commutes.
Let $\cs{F}$ be the virtural Weil sheaf on the group scheme $G= \Gr^{\OK}(\underline{G})$ obtained by pullback along $G \to G_m$.
Then 
\[
\trFrob{\cs{F}} = \trace(\,^\circ \rho),
\] 
as desired.
\end{proof}
